### 2015-02-28 icfp15-servo/rust.tex
Text cleanups
De-LaTeX quotes in code. Soften some of the module-related statements.

Rust is a statically typed systems programming language most heavily inspired by the C
and ML families of languages~\cite{RUST}.
Like the C family of languages, it provides the developer fine control over memory layout
and predictable performance.
Unlike with the C family Rust programs are \emph{memory safe} by default,
only allowing unsafe operations in specially-delineated blocks.

@@

  Thread::spawn(move || {
    *data = *data + 1;
  });
  // error: accessing moved value
  print!(``{}'', data);
}
\end{lstlisting}
  \caption{Code that will not compile because it attempts to access mutable state from two threads.}
  \label{fig:bad-concurrency}
\end{figure}

@@

  Thread::scoped(|| {
    println!(``{}'', data);
  });
  print!(``{}'', data);
}
\end{lstlisting}
  \caption{Safely reading immutable state from two threads.}
  \label{fig:shared-concurrency}
\end{figure}

@@

  print!(``{}'', *data.lock().unwrap());
}
\end{lstlisting}
  \caption{Safely mutating state from two threads.}
  \label{fig:shared-mutable-concurrency}
\end{figure}

>>>>

Rust is a statically typed systems programming language most heavily inspired by the C
and ML families of languages~\cite{RUST}.
Like the C family of languages, it provides the developer fine control over memory layout
and predictable performance.
Unlike C programs, Rust programs are \emph{memory safe} by default,
only allowing unsafe operations in specially-delineated blocks.

@@

  Thread::spawn(move || {
    *data = *data + 1;
  });
  // error: accessing moved value
  print!("{}", data);
}
\end{lstlisting}
  \caption{Code that will not compile because it attempts to access mutable state from two threads.}
  \label{fig:bad-concurrency}
\end{figure}

@@

  Thread::scoped(|| {
    println!("{}", data);
  });
  print!("{}", data);
}
\end{lstlisting}
  \caption{Safely reading immutable state from two threads.}
  \label{fig:shared-concurrency}
\end{figure}

@@

  print!("{}", *data.lock().unwrap());
}
\end{lstlisting}
  \caption{Safely mutating state from two threads.}
  \label{fig:shared-mutable-concurrency}
\end{figure}

### 2015-02-28 icfp15-servo/servo.tex
De-LaTeX quotes in code. Soften some of the module-related statements.

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.
Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied against, providing predictable output code to developers at the cost of some code duplication.
This approach, combined with whole-program compilation~\cite{weeks:whole-program-mlton}, is used in Rust.
This strategy has a negative impact on compilation speed but has resulted in Rust code that easily matches the
sequential speed of its \Cplusplus{} analog without requiring the Servo developers to become compiler experts.

>>>>

\subsection{Compilation strategy}
Many statically typed implementations of polymorphic languages such as Standard ML of New Jersey~\cite{SMLNJ} and
\ocaml{}~\cite{ocaml-manual-3.0} have used a compilation strategy that optimizes representations of data types when
polymorphic code is monomorphically used, but defaults to a less efficient style otherwise, in order to share
code~\cite{ocaml-repr}.
This strategy reduces code size, but leads to unpredictable performance and code, as changes to a codebase that
either add a new instantiation of a polymorphic function at a given type or, in a modular compilation setting, that
expose a polymorphic function externally, can change the performance of code that is not local to the change being
made.

Monomorphization, as in MLton~\cite{mlton-compiler}, instead instantiates each polymorphic code block at each of the types
it is applied against, providing predictable output code to developers at the cost of some code duplication.
This strategy is used by virtually all \Cplusplus{} compilers to implement templates, so it is proven and well-known
within systems programming.
Rust also follows this approach, although it improves on the ergonomics of \Cplusplus{} templates
by embedding serialized generic function ASTs within ``compiled'' binaries.

Rust also chooses a fairly large default compilation unit size. A Rust \textit{crate} is subject to whole-program
compilation~\cite{weeks:whole-program-mlton}, and is optimized as a unit. A crate may comprise hundreds of modules,
which provide namespacing and abstraction. Module dependencies within a crate are allowed to be cyclic.

The large compilation unit size slows down compilation and especially diminishes the ability to build code in parallel.
However, it has enabled us to write Rust code that easily matches the sequential speed of its \Cplusplus{} analog, without requiring the Servo developers to become compiler experts.
Servo contains about 600 modules within 20 crates.
